Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-03-15 paper/tex/results.tex
consistent messaging around figure 1

\section{Results}
\label{sec:results}
%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou)

\subsection{Ensembles of photo-\mathinhead{$z$}{z} interim posteriors}

\begin{figure*}
%\centering
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.5\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
\caption{Four illustrative examples of individual p(z) distributions produced by the codes.  The red vertical line represents the true redshift.  Examples are chosen with common features seen in PDFs: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).  Codes show varying amounts of small-scale structure in their reconstruction of the posterior distribution.  We see varying responses from the codes in the presence of color degeneracies and photometric errors, resulting in narrow and broad unimodal, bimodal, and multi-modal $p(z)$ curves.} \label{fig:pz_examples}
\end{figure*}

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{fig/QQplot_11codes.jpg}
%\caption{The QQ plots produced for every photo-$z$ code.  \aim{Consider making this a plot of QQ minus that diagonal for better contrast between methods.}\red{The PIT histogram is essentially this, which is one of the main reasons to include both.  QQ is a fairly standard form, so plotting in this fashion rather than the difference seems better to me, given that we also ahve the PIT histograms. --SJS}} \label{fig:qq}
%\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{Summary plots for all twelve photo-$z$ codes illustrating performance for the interim posterior statistics. The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).  The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.  The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.  A variety of behaviors are evident, and specific details are discussed in the text.}
\label{fig:pitqq}
\end{figure*}

Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our twelve photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.
Several features are obvious even in these illustrative examples.
\textsc{ANNz2, METAPhoR, CMNN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  For example, in \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is due to the finite number of neighbours, and is not unexpected.
\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of $\delta z = 0.01$ for each PDF.
A piecewise constant grid was a natural choice for some photo-$z$ methods, for instance most template-based codes compute likelihoods on a fixed grid.  %In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
%Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.  We will discuss PDF storage in Section~\ref{sec:discussion}.
%Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
For faint galaxies, this redshift resolution is sufficient to capture the shape of $p(z)$ for the majority of the test sample, where photometric errors on the faint galaxies lead to somewhat broad peaks in the redshift posterior.
However, as can be seen in e.~g.~the top left panel of Fig.~\ref{fig:pz_examples}, for bright galaxies with narrow $p(z)$ the grid spacing of $\delta z = 0.01$ is not always sufficient to resolve the peak.
This is consistent with the results described in \citet[]{Malz:qp}, who find that quantiles (and, to a lesser degree, samples) often outperform gridded $p(z)$, particularly for bright objects and in the presence of harsher storage constraints.
With a full 200 numbers to capture the information of each photo-$z$ PDF, any parametrization will perform adequately, but other storage parametrizations and limits on storage resources may be considered in future work.
%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.  However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
We will discuss this further in Section~\ref{sec:discussion}.
% \red{someone review this statement to make sure that I'm saying this correctly!}

>>>>

\section{Results}
\label{sec:results}
%(Sam Schmidt, Bryce Kalmbach, Johann Cohen Tanugi, Rongpu Zhou, Alex Malz)

We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
Fig.~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.

\begin{figure*}
%\centering
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_261931_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_471167_noseaborn_crop.jpg}\\
\includegraphics[width=0.49\textwidth]{fig/pz_12codes_713178_noseaborn_crop.jpg}\includegraphics[width=0.49\textwidth]{fig/pz_12codes_982747_noseaborn_crop.jpg}
\caption{The individual \pzpdf s (blue) distributions produced by the twelve codes (small panels) on four exemplary galaxies' photometry (large panels) with different true redshifts (red).
The \pzpdf s of all codes share some features for the example galaxies due to physical color degeneracies and photometric errors: tight unimodal $p(z)$ (upper left), broad unimodal $p(z)$ (upper right), bimodal $p(z)$ (lower right), and complex/multimodal $p(z)$ (lower left).
The diverse algorithms and implementations induce differences in small-scale structure and sensitivity to physical systematics.}
\label{fig:pz_examples}
\end{figure*}

The most striking differences between codes are due to small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
The $\rm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a smoking gun for overfitting of the training set, a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question.
The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.
%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
%In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
%Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.  We will discuss PDF storage in Section~\ref{sec:discussion}.
%Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.  However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
% We will discuss this further in Section~\ref{sec:discussion}.
% \red{someone review this statement to make sure that I'm saying this correctly!}

\subsection{Ensembles of \pzpdf s}

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of $0$ and $1$, and asymmetry indicates systematic bias, a form of model misspecification.}
\label{fig:pitqq}
\end{figure*}
